Minimizing Time Required For Graduation

### 2010-12-11 paper/paper.tex
added new section describing quiz section handling and two new graphics

\section{Problem Description} To make our examples concrete we will focus on the
classes offered within the Math department, although our algorithm and model can
be extended to include over departments.

@@

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course is identified by its department name and
number, for example {\it MATH309}. Each course is also worth a specific amount
of credits. \item We refer to a specific instance of a course as a {\it course
offering}. Each course offering additionally has associated with it the time of
the day at which the course meets, the quarter it is being offered, and the
section number of the offering. Throughout this paper, we will refer to quarters
by their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter
Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter.
\end{itemize}

@@

\subsection{Initial Model and Tractability} With a static schedule it would be
possible to plan an entire set of quarter by quarter schedules from start to
finish. Unfortunately we have neither the computational resources nor the luxury
of a static schedule. Some classes are offered every quarter others such as
senior electives are not. Furthermore for some classes it cannot be predicted
more than one quarter in advance when a senior elective will be offered. Even
given perfect information, where we know the possible class schedules for every
quarter we face another problem; finding the best possible schedule reduces to
enumerating possible class schedules for every quarter until the graduation
requirements are satisfied.  Given that we cannot predict in advance the
conflicts between classes, we cannot even determine the branching factor of the
enumeration tree nor its depth. Since we can only determine one quarter in
advance, this suggests approaching the problem by using heuristics to pick the
best possible classes each quarter.

@@

\subsubsection{Procedure} We present the procedure for populating the set $L$ as
follows: \begin{enumerate} \item For each rule $r \in R$ we iterate over each
proposition $p$ in the rule. Recall that $p$ represents a course, for example
$MATH125$. We then check if $p \in T$.  If $p \in T$ then we say $p$ evaluates
to true; otherwise, $p$ evaluates to false.  Using rules of logic we evaluate
$r$, if $r$ true, then we add the course associated with it to $L$.  Once this
is complete, $L$ will consist of every course that could be taken based solely
on prerequisites that we have fulfilled.  \item We now prune $L$ by removing all
courses that are not offered in the current quarter. $L$ now only contain
classes that we can take given the current quarter and the classes we have
already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
every course offering in $L$ is a node, and there is an edge between two nodes
$i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
course offering $j$.  \end{enumerate}

\begin{figure} [ht]
    \begin{center}
        \includegraphics[scale=0.35]{conflict_graph_ex}
    \end{center}
    \caption{A Conflict Graph}
    \label{conflictg}
\end{figure}

>>>>

\section{Problem Description} \label{pdesc} To make our examples concrete we
will focus on the classes offered within the Math department, although our
algorithm and model can be extended to include over departments.

@@

\section{Model Elements} The solve the problem outlined in section \ref{pdesc}
we formalized elements of the academic quarter. Some of the hardest elements to
model were representing prerequisites for a course, how to handle quiz sections,
and simply choosing appropriate sets to contain information about a student's
academic progress.

\subsection{Classes and Quiz Sections} Another modeling difficulty that we
initially ignore was how to represent quiz sections for a class. Each time
a course is offered is may be offered as multiple classes. For instance, during
$F$ $MATH307$ is offered as 7 different sections all teaching the same material
during different time periods. Within a specific quarter schedule line number,
SLNs uniquely identify courses allowing us to differentiate between classes with
the same name. Continuing the above example, the SLNs of the 7 $MATH307$
instances may be 15432, 15433, 15434...

Our peer review team commented that we had not accounted for courses that
required concurrent quiz sections. First we note that the quiz sections for
a class never occur during the same time periods as the class itself (otherwise
students would not be able to attend the quiz section). Quiz sections also have
unique SLNs, to include a required quiz section we include the quiz section SLN
with the SLN for the associated class. Finally, courses such at $MATH125$ can
have multiple sections, each with several quiz sections of which only one need
be taken. We found the easiest way to represent this situation was to iterate
over the quiz sections for each class section and generate a puesdo class for
each pair of class section and quiz section that occured during the time periods
for both quiz and main sections. Fig. \ref{quizzes} contains one section
$MATH125$ being split across 3 quiz sections.
\begin{figure} [ht] 
    \begin{center}
        \includegraphics[scale=0.4]{quiz_sections} 
    \end{center}
    \caption{Splitting up MATH125} 
    \label{quizzes} 
\end{figure}

\subsection{Definitions} We will now formalize some terms used in our paper.
\begin{itemize} \item A {\it course} refers to any course offered by the
University of Washington. A course is identified by its department name and
number, for example {\it MATH309}. Each course is also worth a specific amount
of credits. \item We refer to a specific instance of a course as a {\it class}.
Each class is has associated with it the time of the day at which it  meets, the
quarter it is being offered, and the section number of the offering.  Throughout
this paper, we will refer to quarters by their time schedule abbreviates: {\it
F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter,
and {\it S} for Summer Quarter. \end{itemize}

@@

\section{Tractability} With a static schedule it would be
possible to plan an entire set of quarter by quarter schedules from start to
finish. Unfortunately we have neither the computational resources nor the luxury
of a static schedule. Some classes are offered every quarter others such as
senior electives are not. Furthermore for some classes it cannot be predicted
more than one quarter in advance when a senior elective will be offered. Even
given perfect information, where we know the possible class schedules for every
quarter we face another problem; finding the best possible schedule reduces to
enumerating possible class schedules for every quarter until the graduation
requirements are satisfied.  Given that we cannot predict in advance the
conflicts between classes, we cannot even determine the branching factor of the
enumeration tree nor its depth. Since we can only determine one quarter in
advance, this suggests approaching the problem by using heuristics to pick the
best possible classes each quarter.

@@

\subsubsection{Procedure} We present the procedure for populating the set $L$ as
follows: \begin{enumerate} \item For each rule $r \in R$ we iterate over each
proposition $p$ in the rule. Recall that $p$ represents a course, for example
$MATH125$. We then check if $p \in T$.  If $p \in T$ then we say $p$ evaluates
to true; otherwise, $p$ evaluates to false.  Using rules of logic we evaluate
$r$, if $r$ true, then we add the course associated with it to $L$.  Once this
is complete, $L$ will consist of every course that could be taken based solely
on prerequisites that we have fulfilled.  \item We now prune $L$ by removing all
courses that are not offered in the current quarter. $L$ now only contain
classes that we can take given the current quarter and the classes we have
already taken.  \item We now construct a conflict graph, $G_{conflict}$, where
every course offering in $L$ is a node, and there is an edge between two nodes
$i$ and $j$ if and only if the time of course offering $i$ overlaps the time of
course offering $j$. An example conflict graph for one time period is given in
Fig. \ref{conflictg}. \end{enumerate} 
\begin{figure} [ht] 
    \begin{center}
        \includegraphics[scale=0.35]{conflict_graph_ex} 
    \end{center} 
    \caption{A Conflict Graph} 
    \label{conflictg} 
\end{figure}
